---
title: "Luminex_DataAnalysis"
author: "Elise Viox"
date: "November 15th, 2018"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```
```

##R Markdown
#Analyze Luminex Data
##Generate Practice Luminex Data
Multiple practice tables were created with data similar to what we might obtain from running Luminex and combined into one table.

###Create Datatable 1
```{r Create Data Table #1}
#Create "fake1" datatable
library(knitr)
library(kableExtra)
library(data.table)
donor<-c(rep("A_SMpos", 6), rep("B_SMneg",6), rep("C_SMneg",6), rep("D_SMneg",6), rep("E_SMpos",6), rep("F_SMpos",6), rep("G_SMneg",6), rep("H_SMneg",6), rep("I_SMneg",6), rep("J_SMpos",6), rep("K_SMpos",6), rep("L_SMpos",6), rep("M_SMpos", 6))
stim<-rep(c("un","w","p","s","sw","sb"),3)
ifng<-rnorm(70, 2)
tnfa<-rnorm(70,10)
IL4<-rnorm(70,5)
IL5<-rnorm(70,7)
IL10<-rnorm(70,13)
IL13<-rnorm(70,19)
IL17<-rnorm(70,17)
IL21<-rnorm(70,11)
IL22<-rnorm(70,3)

fake1<-as.data.table(cbind(donor,stim,ifng,tnfa,IL4, IL5, IL10, IL13, IL17, IL21, IL22))
fake1$ifng=as.numeric(as.character(fake1$ifng))
fake1$tnfa=as.numeric(as.character(fake1$tnfa))
fake1$IL4=as.numeric(as.character(fake1$IL4))
fake1$IL5=as.numeric(as.character(fake1$IL5))
fake1$IL10=as.numeric(as.character(fake1$IL10))
fake1$IL13=as.numeric(as.character(fake1$IL13))
fake1$IL17=as.numeric(as.character(fake1$IL17))
fake1$IL21=as.numeric(as.character(fake1$IL21))
fake1$IL22=as.numeric(as.character(fake1$IL22))

#Print "fake1" datatable 
library(knitr)
kable(fake1) %>% kable_styling(latex_options="scale_down")

#Write "fake1" csv
setwd("/Users/eviox/Documents/Emory_IMP/Rotations/Day_Lab/Luminex/FakeData/")
write.csv(fake1,"fake1.csv")
```

###Create Datatable 2
```{r Create Data Table #2}
#Create "fake2" datatable
library(knitr)
library(kableExtra)
library(data.table)
donor<-c(rep("N_SMpos",6), rep("O_SMpos",6), rep("P_SMpos",6), rep("Q_SMneg",6), rep("R_SMpos",6), rep("S_SMneg",6), rep("T_SMneg",6), rep("U_SMneg",6), rep("V_SMneg",6), rep("W_SMpos",6), rep("X_SMpos",6), rep("Y_SMneg", 6), rep("Z_SMpos", 6))
stim<-rep(c("un","w","p","s","sw","sb"),3)
ifng<-rnorm(70, 2)
tnfa<-rnorm(70,10)
IL4<-rnorm(70,5)
IL5<-rnorm(70,7)
IL10<-rnorm(70,13)
IL13<-rnorm(70,19)
IL17<-rnorm(70,17)
IL21<-rnorm(70,11)
IL22<-rnorm(70,3)

fake2<-as.data.table(cbind(donor,stim,ifng,tnfa,IL4, IL5, IL10, IL13, IL17, IL21, IL22))
fake2$ifng=as.numeric(as.character(fake1$ifng))
fake2$tnfa=as.numeric(as.character(fake1$tnfa))
fake2$IL4=as.numeric(as.character(fake1$IL4))
fake2$IL5=as.numeric(as.character(fake1$IL5))
fake2$IL10=as.numeric(as.character(fake1$IL10))
fake2$IL13=as.numeric(as.character(fake1$IL13))
fake2$IL17=as.numeric(as.character(fake1$IL17))
fake2$IL21=as.numeric(as.character(fake1$IL21))
fake2$IL22=as.numeric(as.character(fake1$IL22))

#Print "fake2" datatable 
library(knitr)
kable(fake2) %>% kable_styling(latex_options="scale_down")

#Write "fake2" csv
setwd("/Users/eviox/Documents/Emory_IMP/Rotations/Day_Lab/Luminex/FakeData/")
write.csv(fake2,"fake2.csv")
```


###Combine Datatables
```{r Combine fake datatables read in from folder}
library(data.table)
library(kableExtra)
library(knitr)

#Identify directory where files of interest are located
file_names <- dir("/Users/eviox/Documents/Emory_IMP/Rotations/Day_Lab/Luminex/FakeData/")

#Set directory for new file to be stored
setwd("/Users/eviox/Documents/Emory_IMP/Rotations/Day_Lab/Luminex/FakeData/")

#Combine all files in the file_names directory into one file
fakecombined <- do.call(rbind,lapply(file_names,read.csv))
fakecombined$X <- NULL

kable(fakecombined) %>% kable_styling(latex_options="scale_down")
```

##Subtract out unstimulated values
I wrote a function called "subtractun"" that pulls the unstimulated value for each cytokine from each donor and subtracts it from the respective stim values for the respective cytokine.

The function uses the 1) split, 2) apply, 3)combine sequence to 1) generate data tables for each individual donor, 2) apply the subtraction of the unstim to the respective cytokines for those donors, and 3) take these newly calculated values for individual donors and combine them into a data table containing all donor values.

The subtractun function will be applied only to the numeric columns of a datatable as specified by the "numeric.only" function. In the case of luminex data, this ensures that the function will only be applied to cytokine data. 

###Split table by donor
```{r Split Table}
library(knitr)
library(kableExtra)
#Redefine donor (since previously defined for generating fake data)
donor<-fakecombined$donor
fakesplit<- split(fakecombined, donor)
kable(fakesplit$A) %>% kable_styling(latex_options="scale_down")
kable(fakesplit$B) %>% kable_styling(latex_options="scale_down")
kable(fakesplit$C) %>% kable_styling(latex_options="scale_down")
kable(fakesplit$D) %>% kable_styling(latex_options="scale_down")
kable(fakesplit$E) %>% kable_styling(latex_options="scale_down")
kable(fakesplit$F) %>% kable_styling(latex_options="scale_down")
```


###Create function for subtracting out unstimulated
```{r Create Split, Apply, Combine Function}
subtractun<-function (datatable) {
  #Split datatable by donor    
    y<- split(datatable, donor)
  #Write function that will select numeric columns of datatable
    numeric.only <- function(X,...){
      returnCols <- names(X)
      a<-sapply(X, is.numeric)
      print(returnCols[a == "TRUE"])
      }
  #Apply numeric.only function to datatable
    for (z in numeric.only(datatable)){
  #Subtract out unstim value from respective stim cytokine values
    newcolumn<-unlist(lapply(y,function(g)
      (g[,z]- as.matrix(subset(g[,z], g$stim=="un"))[1,1])
      ))
  #Create new datatable with unstim substractions applied to all donors 
    datatable[,z] <- newcolumn}
    datatable
      }  
```

###Apply function for subtracting out unstimulated to data table of interest
```{r Apply function to data table of interest}
#Apply subtractun function to datatable of interest
newfake<-subtractun(fakecombined)
newfake[newfake<0] <- 0
library(knitr)
library(kableExtra)
kable(newfake) %>% kable_styling(latex_options="scale_down")
library(dplyr)
library(tidyr)
newfake<-newfake %>%
  separate(donor, c("donor", "Status"), "_")
```

##Reshape data table so that each cytokine name is assigned to row entries instead of having its own column
I applied the gather function to make the data table format compatible for the next set of functions.
```{r Gather Datatable}
library(tidyr)
library(knitr)

gathereddata<-gather(newfake, "Cytokine", "Concentration", c("ifng", "tnfa", "IL4", "IL5", "IL10", "IL13", "IL17", "IL21", "IL22"))

kable(gathereddata)
```

##Calcualte p values for comparisons between SMneg and SMpos for different cytokines for different treatments
```{r P vals}
#Function for Wilcox rank sum 2-sample test
##Subset data by condition
##Split subsetted data by cytokine
##Run Wilcox rank sum 2-sample test by Schisto status
library(knitr)
pvals<-function(data,column,condition){
SubsettedData<-subset(data, data[,column]== condition)
x<-split(data,list(data$Cytokine))
A<-lapply(x, function(g) wilcox.test(g$Concentration~g[,"Status"]))
pvalslist<-c(A$infg$p.value,A$tnfa$p.value,A$IL4$p.value,A$IL5$p.value,A$IL10$p.value,A$IL13$p.value,A$IL17$p.value,A$IL21$p.value, A$IL22$p.value)
Analytes<-c("ifng","tnfa","IL4","IL5", "IL10","IL13", "IL17", "IL21", "IL22")
Pvalues<-round(pvalslist, 4)
table<-(cbind(Analytes, Pvalues))
}

#Apply function to desired stim condition of gathered data table
WilcoxTestTable_WCL<-pvals(gathereddata,"stim","w")
kable(WilcoxTestTable_WCL) 

#Future direction: Adjust the p values to account for multiple comparisons
##hochberg_adj_pvals<-round(p.adjust(pvalslist, method="hochberg"), 4)
```

#Perform Hierarchical Clustering Analysis
##Generate Heatmap for each stim condition
I created one heatmap per stim condition showing the differences in concentration of cytokines between patients. 
```{r Heatmap Analysis}
library(ggplot2)
library(stats)

#Create general heatmap function
heatmapfunction<-function(data,column,condition){
SubsettedData<-subset(data, data[,column]== condition)
heatmap <- ggplot(data = SubsettedData, mapping = aes(x = donor,
              y = Cytokine,
              fill = Concentration)) +
              geom_tile() +
              xlab(label = "Sample") +
              ggtitle(label = condition) +
              facet_grid(~Status, switch = "x", scales = "free_x", space = "free_x")
}
#Apply function to stimulation condition of interest in indicated datatable
SEB<-heatmapfunction(gathereddata,"stim","sb")
SEA<-heatmapfunction(gathereddata,"stim","s")
SWAP<-heatmapfunction(gathereddata,"stim","sw")
WCL<-heatmapfunction(gathereddata,"stim","w")
PeptidePool<-heatmapfunction(gathereddata,"stim","p")

SEB
SEA
SWAP
WCL
PeptidePool
```

#Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

##Generate Dendrogram for each stim condition
```{r Create Dendrogram}

```

##Merge Heatmap and Dendrogram
```{r Merge Heatmap and Dendrogram}

```

##Order heatmap according to position in dendrogram to generate hierarchically clustered heatmap
```{r Order the heatmap according to position in the cluster}

```

